Supplementary Material for _A Guide to Pre-processing High-Frequency Animal Tracking Data_

### 2021-06-11 supplement/05_calibration_data.Rmd
edits per reviewer comments

`data_frame[, c("col_a", "col_b) := list(values_a, values_b)]`

@@

First we access the data from a local file using the `data.table` package [@dowle2020].
We look at the first few rows, using `head()`.
We then visualise the raw data.

@@

We first save a copy of the data, so that we can plot the raw data with the cleaned data plotted over it for comparison.

@@

We then plot the result of filtering, with the excluded point in black, and the points that are retained in green.

@@

@@

Next we thin the data to demonstrate thinning by median smoothing.
Following this, we plot the median smooth and thinning by aggregation.

@@

# remove columns we don't need
data <- data[, setdiff(
  colnames(data),
  c("tID", "Timestamp", "id", "TIME", "UTCtime")
),
with = FALSE
]

@@

# now get times 10 mins before and after
data_res_summary[, `:=`(
  t_min = t_median - (10 * 60),
  t_max = t_median + (10 * 60)
)]

# make a list of positions 10min before and after
wp_data <- mapply(function(l, u, mx, my) {
  tmp_data <- data_unproc[inrange(time, l, u)]
  tmp_data[, distance := sqrt((mx - x)^2 + (my - y)^2)]

  # keep within 50
  tmp_data <- tmp_data[distance <= 50, ]

  # get duration
  return(diff(range(tmp_data$time)))
}, data_res_summary$t_min, data_res_summary$t_max,
data_res_summary$x_median, data_res_summary$y_median,
SIMPLIFY = TRUE
)
```

@@

First, we filter data where we know the animal (or in this case, the human-carried tag) spent some time at or near a position, as this is the first step to identify residence patches. 
One way of doing this is to filter out positions with speeds above which the tag (ideally on an animal) is likely to be in transit.
Rather than filtering on instantaneous speed estimates, filtering on a median smoothed speed estimate is more reliable.

>>>>

`data_frame[, c("col_a", "col_b") := list(values_a, values_b)]`

@@

First we access the data from a local file using the `data.table` package [@dowle2020].

In all, we aim to keep three versions of the data: (1) `data_raw`, the entirely unprocessed data, (2) `data`, the working version, and (3) `data_unproc`, data that has been partially processed, but which is one step behind `data`. This allows us to better illustrate the pre-processing steps, and prevents us from irreversibly modifying our data --- at best, we would have to re-run many pre-processing steps, and at worse, we might overwrite the data on disk.

We look at the first few rows, using `head()`. We then visualise the raw data.

@@

We first save a copy of the data, so that we can plot the unprocessed data with the cleaned data plotted over it for comparison. Here, `data_unproc`, `data`, and `data_raw` are still the same, since no pre-processing steps have been applied yet.

@@

We then plot the result of filtering, with the excluded point in black, and the points that are retained in green. After this stage, `data` is filtered and 'ahead' of `data_raw` and `data_unproc`, which are still the same. This pattern will repeat throughout this material.

@@

Here, `data_unproc` moves 'ahead' of `data_raw`, and holds the data filtered by a bounding box --- `data` is also moving ahead, and will be filtered on speed.

@@

Next we thin the data by aggregation to demonstrate thinning after median smoothing.
Following this, we plot the median smooth and thinning by aggregation.

@@

# remove columns we don't need
data <- data[, !c("tID", "Timestamp", "id", "TIME", "UTCtime")]

@@

# now get times 10 mins before and after
data_res_summary[, c("t_min", "t_max") := list(
  t_median - (10 * 60),
  t_median + (10 * 60)
)]

# manually get the duration of the stops
wp_data <- mapply(function(l, u, mx, my) {

  # first select all data whose timestamp is between
  # the upper and lower bounds of the stop (l = lower, u = upper)
  tmp_data <- data_unproc[inrange(time, l, u), ]

  # calculate the distance between the positions selected above
  # and the median X and Y coordinates of the stop (centroid)
  tmp_data[, distance := sqrt((mx - x)^2 + (my - y)^2)]

  # keep positions that are within 50m of the centroid
  tmp_data <- tmp_data[distance <= 50, ]

  # get the duration of the stop as the difference between
  # the minimum and maximum times of the positions retained above
  return(diff(range(tmp_data$time)))
}, data_res_summary$t_min, data_res_summary$t_max,
   data_res_summary$x_median, data_res_summary$y_median,

   # this specifies that a vector, rather than a list, is returned
   SIMPLIFY = TRUE
)
```

@@

First, we filter data where we know the animal (or in this case, the human-carried tag) spent some time at or near a position, as this is the first step to identify residence patches. 
One way of doing this is by filtering out positions with speeds above which the tag (ideally on an animal) is likely to be in transit.
Rather than filtering on instantaneous speed estimates, filtering on a median smoothed speed estimate is more reliable.
